---
title: "BankChurn"
author: "John Lee"
date: "4/18/2021"
output:
  pdf_document: default
  word_document: default
---
##Data Processing
```{r}
library(readr)
library(tidyverse)
library(patchwork)
library(caret)
library(vcd)
library(gridExtra)
library(knitr)
library(corrplot)
library(scales)
library(lme4)
library(DMwR2)
library(InformationValue)
library(ROCR)
library(rpart)
library(randomForest)
library(MASS)
library(ggmosaic)
library(e1071)
library(ranger)
library(penalized)
library(rpart.plot)
library(ggcorrplot)
library(caTools)
library(plyr)
library(corrplot)
library(ggplot2)
library(party)
library(vip)
```


```{r}
churn <- read_csv("C:/Users/14044/Desktop/Wake Forest/7095/Excel Files/BankChurners_clean.csv");
str(churn)
```

##Data Cleaning
#---Change Attrition_Flag encoding: Existing Customer = 0, Attrition Customer = 1
```{r}
churn <- churn %>% mutate(Attrition_Flag = recode(Attrition_Flag, "Attrited Customer" = 1, "Existing Customer" = 0))

```
#---Factor conversion
```{r}

churn$Attrition_Flag <- as_factor(churn$Attrition_Flag)

churn$Gender <- as_factor(churn$Gender)

churn$Education_Level <- as.factor(churn$Education_Level)
                                   
churn$Marital_Status <- as.factor(churn$Marital_Status)

churn$Income_Category <- as.factor(churn$Income_Category)

churn$Card_Category <- as.factor(churn$Card_Category)
str(churn)
```
#---Check to see if there's any missing data
```{r}
sapply(churn, function(x) sum(is.na(x)))
```

##Exploratory Data Analysis
```{r}
summary(churn)
```

```{r}
ggplot(churn, aes(Attrition_Flag, fill = Attrition_Flag)) +
  geom_bar() +
  theme(legend.position = 'none')
```

```{r}
table(churn$Attrition_Flag)
```

```{r}
round(prop.table(table(churn$Attrition_Flag)),3)
```

#---Continuous Variable Distribution--- 
```{r}
churn %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot() +
  geom_histogram(mapping = aes(x=value,fill=key), color="black") +
  facet_wrap(~ key, scales = "free") +
  theme_minimal() +
  theme(legend.position = 'none')
```

#---Find correlations between columns---
```{r}
numeric.var <- sapply(churn, is.numeric)
corr.matrix <- cor(churn[,numeric.var])
corrplot(corr.matrix, main="\n\nCorrelation Plot for Numerical Variables", method="number")
```

#---Remove high correlating numerical variables---
```{r}
churn <- subset(churn, select = -c(Months_on_book, Total_Trans_Ct, Total_Revolving_Bal))
```

#---Continuous Variable Exploration
```{r}
#---Age---
age_hist <- ggplot(churn, aes(x=Customer_Age, fill = Attrition_Flag)) +
  geom_histogram(binwidth = 5) + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,100, by=10))

age_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Customer_Age, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

age_hist 
age_boxplot
```


```{r}
#---Total Revovling Balance---
credit_limit_hist <- ggplot(churn, aes(x=Credit_Limit, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

credit_limit_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Credit_Limit, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

credit_limit_hist
credit_limit_boxplot
```


```{r}
#---Dependent_count---
Dep_count_hist <- ggplot(churn, aes(x=Dependent_count, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,10, by=1))

Dep_count_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Dependent_count, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Dep_count_hist
Dep_count_boxplot

```


```{r}
#---Total Relationship Count
Tot_relat_hist <- ggplot(churn, aes(x=Total_Relationship_Count, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,10, by=1))

Tot_relat_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Total_Relationship_Count, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Tot_relat_hist
Tot_relat_boxplot

```


```{r}
#---Months Inactive 12 mon
Months_Inactive_hist <- ggplot(churn, aes(x=Months_Inactive_12_mon, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,10, by=1))

Months_Inactive_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Months_Inactive_12_mon, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Months_Inactive_hist
Months_Inactive_boxplot

```


```{r}
#---Contacts_Count_12_mon---
Contract_count_hist <- ggplot(churn, aes(x=Contacts_Count_12_mon, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,10, by=1))

Contract_count_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Contacts_Count_12_mon, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Contract_count_hist
Contract_count_boxplot

```


```{r}
#---Avg_Open_To_Buy---
Avg_buy_hist <- ggplot(churn, aes(x=Avg_Open_To_Buy, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

Avg_buy_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Avg_Open_To_Buy, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Avg_buy_hist
Avg_buy_boxplot
```


```{r}
#---Total_Amt_Chng_Q4_Q1---
Total_Amt_Chng_Q4_Q1_hist <- ggplot(churn, aes(x=Total_Amt_Chng_Q4_Q1, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

Total_Amt_Chng_Q4_Q1_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Total_Amt_Chng_Q4_Q1, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Total_Amt_Chng_Q4_Q1_hist
Total_Amt_Chng_Q4_Q1_boxplot
```

```{r}
#---Total_Trans_Amt---
Total_Trans_Amt_hist <- ggplot(churn, aes(x=Total_Trans_Amt, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

Total_Trans_Amt_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Total_Trans_Amt, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Total_Trans_Amt_hist
Total_Trans_Amt_boxplot
```


```{r}
#---Total_Ct_Chng_Q4_Q1---
Total_Ct_Chng_Q4_Q1_hist <- ggplot(churn, aes(x=Total_Ct_Chng_Q4_Q1, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

Total_Ct_Chng_Q4_Q1_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Total_Ct_Chng_Q4_Q1, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Total_Ct_Chng_Q4_Q1_hist
Total_Ct_Chng_Q4_Q1_boxplot
```


```{r}
#---Avg_Utilization_Ratio---
Avg_Utilization_Ratio_hist <- ggplot(churn, aes(x=Avg_Utilization_Ratio, fill = Attrition_Flag)) +
  geom_histogram() + 
  theme_minimal() + 
  scale_x_continuous(breaks = seq(0,50000, by=10000))

Avg_Utilization_Ratio_boxplot <- ggplot(churn, aes(x= Attrition_Flag, y = Avg_Utilization_Ratio, fill = Attrition_Flag)) +
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = 'none')

Avg_Utilization_Ratio_hist
Avg_Utilization_Ratio_boxplot
```

##---Categorical Variable Distribution---
```{r}
p1 <- ggplot(churn, aes(x=Gender)) + ggtitle("Gender") + xlab("Gender") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p2 <- ggplot(churn, aes(x=Education_Level)) + ggtitle("Education Level") + xlab("Education Level") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p3 <- ggplot(churn, aes(x=Marital_Status)) + ggtitle("Marital Status") + xlab("Marital Status") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p4 <- ggplot(churn, aes(x=Income_Category)) + ggtitle("Income Category") + xlab("Income Category") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p5 <- ggplot(churn, aes(x=Card_Category)) + ggtitle("Card Category") + xlab("Card Category") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
grid.arrange(p1, p2, p3, p4, p5, ncol=2)
```

#---Categorical Variable Exploration---

```{r}
gender_graph <- churn %>%
  dplyr::select(Gender, Attrition_Flag) %>% 
  table(.) %>% 
  as.data.frame() %>% 
  ggplot(.) +
  ggmosaic::geom_mosaic(aes(weight = Freq, x = product(Gender), fill = Attrition_Flag)) +
  ggthemes::theme_tufte() +
  scale_fill_brewer(type = "qual") +
  labs(x = 'Gender')

edc_lvl_graph <- churn %>%
  dplyr::select(Education_Level, Attrition_Flag) %>% 
  table(.) %>% 
  as.data.frame() %>% 
  ggplot(.) +
  ggmosaic::geom_mosaic(aes(weight = Freq, x = product(Education_Level), fill = Attrition_Flag)) +
  ggthemes::theme_tufte() +
  scale_fill_brewer(type = "qual") +
  labs(x = 'Education_level')

Marital_status_graph <- churn %>%
  dplyr::select(Marital_Status, Attrition_Flag) %>% 
  table(.) %>% 
  as.data.frame() %>% 
  ggplot(.) +
  ggmosaic::geom_mosaic(aes(weight = Freq, x = product(Marital_Status), fill = Attrition_Flag)) +
  ggthemes::theme_tufte() +
  scale_fill_brewer(type = "qual") +
  labs(x = 'Marital_Status')

Income_Category_graph <- churn %>%
  dplyr::select(Income_Category, Attrition_Flag) %>% 
  table(.) %>% 
  as.data.frame() %>% 
  ggplot(.) +
  ggmosaic::geom_mosaic(aes(weight = Freq, x = product(Income_Category), fill = Attrition_Flag)) +
  ggthemes::theme_tufte() +
  scale_fill_brewer(type = "qual") +
  labs(x = 'Income_Category')

Card_Category_graph <- churn %>%
  dplyr::select(Card_Category, Attrition_Flag) %>% 
  table(.) %>% 
  as.data.frame() %>% 
  ggplot(.) +
  ggmosaic::geom_mosaic(aes(weight = Freq, x = product(Card_Category), fill = Attrition_Flag)) +
  ggthemes::theme_tufte() +
  scale_fill_brewer(type = "qual") +
  labs(x = 'Card_Category')
  

(gender_graph | edc_lvl_graph) / (Marital_status_graph | Income_Category_graph ) / (Card_Category_graph)
```

#---chi-square test---
```{r}
chi.square <- vector()
p.value <- vector()
cateVar <- churn %>% 
  dplyr::select(-Attrition_Flag) %>% 
  keep(is.factor)

for (i in 1:length(cateVar)) {
 p.value[i] <- chisq.test(churn$Attrition_Flag, unname(unlist(cateVar[i])), correct = FALSE)[3]$p.value
 chi.square[i] <- unname(chisq.test(churn$Attrition_Flag, unname(unlist(cateVar[i])), correct = FALSE)[1]$statistic)
}

chi_sqaure_test <- tibble(variable = names(cateVar)) %>% 
  add_column(chi.square = chi.square) %>% 
  add_column(p.value = p.value)
knitr::kable(chi_sqaure_test)
```

##Predictive Analysis
```{r}
#---split data set to train and test
set.seed(1234)
sample_set <- churn %>%
  pull(.) %>% 
  sample.split(SplitRatio = .7)

bankTrain <- subset(churn, sample_set == TRUE)
bankTest <- subset(churn, sample_set == FALSE)
```

```{r}
#---check distribution of data
round(prop.table(table(churn$Attrition_Flag)),3)
round(prop.table(table(bankTrain$Attrition_Flag)),3)
round(prop.table(table(bankTrain$Attrition_Flag)),3)
```

#---Logistic Regression
```{r}
## Train the model
logit.mod <- glm(Attrition_Flag ~., family = binomial(link = 'logit'), data = bankTrain)

## Look at the result
summary(logit.mod)
```

```{r}
#---predict outcomes against the test data
logit.pred.prob <- predict(logit.mod, bankTest, type = 'response')
logit.pred <- as.factor(ifelse(logit.pred.prob > 0.5, 1, 0))
head(bankTest,10)
head(logit.pred.prob,10)
```

```{r}
#---confusion matrix of logistic regression
caret::confusionMatrix(logit.pred, bankTest$Attrition_Flag, positive = "1")
```

#---Decision Tree
```{r}
ctrl <-
  trainControl(method = "cv", #cross-validation
               number = 10, #10-fold
               selectionFunction = "best")

grid <- 
  expand.grid(
    .cp = seq(from=0.0001, to=0.005, by=0.0001)
  )
set.seed(1234)
tree.mod <-
  train(
    Attrition_Flag ~.,
    data = bankTrain,
    method = "rpart",
    metric = "Kappa",
    trControl = ctrl,
    tuneGrid = grid
  )

tree.mod
```

```{r}
#---predict outcomes against the test data
tree.pred.prob <- predict(tree.mod, bankTest, type = "prob")
tree.pred <- predict(tree.mod, bankTest, type = "raw")
```

```{r}
#---confusion matrix of decision tree
caret::confusionMatrix(tree.pred, bankTest$Attrition_Flag, positive = "1")
```

#---Random Forest Model
```{r}
## Create a control object.
ctrl <- trainControl(method = "cv",
                     number = 10,
                     selectionFunction = "best")

## Grid search based on the available parameters.
grid <- expand.grid(.mtry = c(1:8))

## Random forest model
rf.mod <- 
  train(Attrition_Flag ~.,
        data = bankTrain,
        method = 'rf',
        metric = 'Kappa',
        trControl = ctrl,
        tuneGrid = grid)

rf.mod
```

```{r}
#---make predictions
rf.pred <- predict(rf.mod, bankTest, type = "raw")
rf.pred.prob <- predict(rf.mod, bankTest, type = "prob")
```

```{r}
#---confusion matrix of Random Forest Model
caret::confusionMatrix(rf.pred, bankTest$Attrition_Flag, positive = "1")
```


##Compare the models
```{r}
## Logistic Regression
test <- bankTest$Attrition_Flag
pred <- logit.pred
prob <- logit.pred.prob

# Logistic Regression ROC curve
roc.pred <- prediction(predictions = prob, labels = test)
roc.perf <- performance(roc.pred, measure = "tpr", x.measure = "fpr")
plot(roc.perf, main = "ROC Curve for Bank Churn Prediction Approaches", col = 2, lwd = 2)
abline(a = 0, b = 1, lwd = 3, lty = 2, col = 1)

## Logistic Regression Performance Metrics
accuracy <- mean(test == pred)
precision <- posPredValue(pred, test, positive = "1")
recall <- caret::sensitivity(pred, test, positive = "1")
fmeasure <- (2 * precision * recall)/(precision + recall)
confmat <- caret::confusionMatrix(pred, test, positive = "1")
kappa <- as.numeric(confmat$overall["Kappa"])
auc <- as.numeric(performance(roc.pred, measure = "auc")@y.values)
comparisons <- tibble(approach="Logistic Regression", accuracy = accuracy, fmeasure = fmeasure,kappa = kappa, auc = auc)

## Classification Tree
test <- bankTest$Attrition_Flag
pred <- tree.pred
prob <- tree.pred.prob[,2]

## Classification Tree ROC Curve
roc.pred <- prediction(predictions = prob, labels = test)
roc.perf <- performance(roc.pred, measure = "tpr", x.measure = "fpr")
plot(roc.perf, col=3, lwd = 2, add=TRUE)

## Classification Tree Performance Metrics
accuracy <- mean(test == pred)
precision <- posPredValue(pred, test, positive = "1")
recall <- caret::sensitivity(pred, test, positive = "1")
fmeasure <- (2 * precision * recall)/(precision + recall)
confmat <- caret::confusionMatrix(pred, test, positive = "1")
kappa <- as.numeric(confmat$overall["Kappa"])
auc <- as.numeric(performance(roc.pred, measure = "auc")@y.values)
comparisons <- comparisons %>%
  add_row(approach="Classification Tree", accuracy = accuracy, fmeasure = fmeasure, kappa = kappa, auc = auc) 

## Random Forest
test <- bankTest$Attrition_Flag
pred <- rf.pred
prob <- rf.pred.prob[,2]

## Random Forest ROC Curve
roc.pred <- prediction(predictions = prob, labels = test)
roc.perf <- performance(roc.pred, measure = "tpr", x.measure = "fpr")
plot(roc.perf, col=4, lwd = 2, add=TRUE)

## Random Forest Performance Metrics
accuracy <- mean(test == pred)
precision <- posPredValue(pred, test, positive = "1")
recall <- caret::sensitivity(pred, test, positive = "1")
fmeasure <- (2 * precision * recall)/(precision + recall)
confmat <- caret::confusionMatrix(pred, test, positive = "1")
kappa <- as.numeric(confmat$overall["Kappa"])
auc <- as.numeric(performance(roc.pred, measure = "auc")@y.values)
comparisons <- comparisons %>%
  add_row(approach="Random Forest", accuracy = accuracy, fmeasure = fmeasure, kappa = kappa, auc = auc) 

# Draw ROC legend.
legend(0.6, 0.6, c('Logistic Regression', 'Classification Tree', 'Random Forest'), 2:5)
```

```{r}
#---output comparison
knitr::kable(comparisons)
```

```{r}
#---Logistic Regression Model Features
vip::vip(logit.mod)
```

```{r}
#---Decsion Tree Features
vip::vip(tree.mod)
```

```{r}
#---Random Forest Model Features
vip::vip(rf.mod)
```

```{r}
vip::vip(rf.mod,5)
```

